Why Real-World Subgraph Isomorphism is Easy, \\ and What This Means for Graph Databases

### 2016-09-25 paper/paper.tex
Some more things

What if we alter the edge probabilities for both the pattern graph and the target graph?  In the top
row of \cref{figure:non-induced} we show the satisfiability phase transition for the non-induced
variant, for patterns of order 10, 20 and 30, targets of order 150, and varying pattern (x-axis) and
target (y-axis) edge probabilities. Each axis runs over 101 edge probabilities, from 0 to 1 in steps
of 0.01, except for the VF2 row which uses steps of 0.02. For each of these points, we generate ten
random instances. The colour denotes the proportion of these instances which were found to be
satisfiable.  Inside the orange region, at the bottom right of each plot, every instance is
unsatisfiable---here we are trying to find a dense pattern in a sparse target. In the purple region,
at the top left, every instance is satisfiable---we are looking for a sparse pattern in a dense
target (which is easy, since we only have to preserve adjacency, not non-adjacency). The white band
between the regions shows the location of the phase transition: here, roughly half the instances are
satisfiable. (We discuss the black line below.)

@@

\begin{figure}[t]
    \centering\includegraphics{gen-graph-labels.pdf}
    \caption{On the top row, predicted and actual location of the phase transition for labelled
    non-induced random subgraph isomorphism, with a pattern order of 20, a target order of 100,
    varying pattern (x-axis) and target (y-axis) density, and varying numbers of labels. On the second
    row, the average number of search nodes needed to solve an instance.}\label{figure:labels}
\end{figure}

@@

\subsection{Flawed and Obviously Unsatisfiable Instances}

@@

@@

These beliefs are not entirely unfounded: none of the matching algorithms considered in these papers
will immediately detect if a pattern contains two red vertices, whilst the target graph contains
only one. This kind of flaw \emph{should} be picked up at the top of search by an all-different
propagator \citep{DBLP:conf/aaai/Regin94}.  However, as \citet{DBLP:journals/pvldb/KatsarouNT15}
note, some variation on VF2 \citep{Cordella:2004} is the usual matching algorithm of choice for
graph database systems, although \citeS{DBLP:journals/jacm/Ullmann76}'s algorithm is sometimes
chosen. Other approaches have been considered: for example, \citet{DBLP:journals/pvldb/ShangZLY08}
propose an algorithm which makes use of the frequency of various features to guide search.
\citet{DBLP:journals/pvldb/LeeHKL12} determine experimentally that this technique tends to be very
effective, even on families of graphs for which it was not designed.

Thus we believe it would be unlikely to cause too much astonishment if we suggested that a better
matching algorithm could be dropped in as a black box replacement in graph databases systems to
improve their performance. However, it appears not to have occurred to the graph databases community
that a better subgraph isomorphism algorithm could invalidate the premise underlying the entire
filter / verify approach. Eliminating this filtering step would not merely be a matter of
simplicity: indices take ??  hours to build.

@@

?? The point of this paper is not to present the ultimate big data subgraph matching algorithm.

@@

?? We need optimisations for trivial cases. Lazy forward checking, cheap alternatives to heuristics and
all-different. Presolving. Huge domains.

@@

@@

>>>>

What if we alter the edge probabilities for both the pattern graph and the target graph?  In the top
row of \cref{figure:non-induced} we show the satisfiability phase transition for the non-induced
variant, for patterns of order 10, 20 and 30, targets of order 150, and varying pattern (x-axis) and
target (y-axis) edge probabilities. Each axis runs over 101 edge probabilities, from 0 to 1 in steps
of 0.01. For each of these points, we generate ten random instances. The colour denotes the
proportion of these instances which were found to be satisfiable.  Inside the orange region, at the
bottom right of each plot, every instance is unsatisfiable---here we are trying to find a dense
pattern in a sparse target. In the purple region, at the top left, every instance is
satisfiable---we are looking for a sparse pattern in a dense target (which is easy, since we only
have to preserve adjacency, not non-adjacency). The white band between the regions shows the
location of the phase transition: here, roughly half the instances are satisfiable. (We discuss the
black line below.)

@@

\begin{figure}[t]
    \centering\includegraphics{gen-graph-labels.pdf}
    \caption{On the top row, predicted and actual location of the phase transition for labelled
    non-induced random subgraph isomorphism, with a pattern order of 20, a target order of 150,
    varying pattern (x-axis) and target (y-axis) density, and varying numbers of labels. On
    subsequent rows, the average number of search nodes needed to solve an instance, for three
    different solvers.}\label{figure:labels}
\end{figure}

@@

\subsection{What's Wrong with VF2?}

@@

These beliefs are not entirely unfounded: none of the matching algorithms considered in these papers
will immediately detect if a pattern contains two red vertices, whilst the target graph contains
only one. This kind of flaw \emph{should} be picked up at the top of search by an all-different
propagator \citep{DBLP:conf/aaai/Regin94}.  However, as \citet{DBLP:journals/pvldb/KatsarouNT15}
note, some variation on VF2 \citep{Cordella:2004} is the usual matching algorithm of choice for
graph database systems, although \citeS{DBLP:journals/jacm/Ullmann76}'s algorithm is sometimes
chosen. (Other approaches have been considered: for example, \citet{DBLP:journals/pvldb/ShangZLY08}
propose an algorithm which makes use of the frequency of various features to guide search.
\citet{DBLP:journals/pvldb/LeeHKL12} determine experimentally that this technique tends to be very
effective, even on families of graphs for which it was not designed.) We believe it would be
unlikely to cause too much astonishment if we suggested that a better matching algorithm could be
dropped in as a black box replacement in graph databases systems to improve their performance.
However, it appears not to have occurred to the graph databases community that a better subgraph
isomorphism algorithm could invalidate the premise underlying the entire filter / verify approach.
Eliminating this filtering step would not merely be a matter of simplicity: indices take ??  hours
to build.

@@

@@

?? Presenting the ultimate big data subgraph matching algorithm is outwith the scope of this paper.

@@

?? We need optimisations for trivial cases. Lazy forward checking, cheap alternatives to heuristics and
all-different. Presolving. Huge domains. Ongoing use of VF2 and micro vs millisecond benchmarks.

@@

?? Use of VF2 in other areas.

@@

\clearpage
\appendix

\section{Reproducing These Experiments}
